Realizing temporal graphs from fastest travel times

### 2023-09-18 ISAAC2023/ISAAC23-longversion.tex
ISAC changes

\begin{abstract}
In this paper we initiate the study of the \emph{temporal graph realization} problem with respect to the fastest path durations among its vertices, 
while we focus on periodic temporal graphs. 
Given an $n \times n$ matrix $D$ and a $\Delta \in \mathbb{N}$, the goal is to construct a $\Delta$-periodic temporal graph with $n$ vertices 
such that the duration of a \emph{fastest path} from $v_i$ to $v_j$ is equal to $D_{i,j}$, or to decide that such a temporal graph does not exist. 
The variations of the problem on static graphs has been well studied and understood since the 1960's, and this area of research remains active until nowadays.

@@

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity among all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all these cases, the duration of the temporal paths has not been considered.

@@

We remark that a fastest path between two vertices in a temporal graph can be computed in polynomial time~\cite{xuan_computing_2003,Wu2016Efficient}.
%Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$, in a temporal graph $(G,\lambda)$, to all other vertices,
%that runs in polynomial-time.
Hence, given a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, we can compute in polynomial-time  the matrix $D$
which consists of durations of fastest temporal paths among all pairs of vertices in $(G,\lambda,\Delta)$.

@@

 We start by describing the vertex set of the underlying graph $G$ of $D$.
\begin{itemize}
\item For each variable $x_i$ in $\phi$, we create three variable vertices $x_i, x_i^T, x_i^F$.
\item For each clause $c$ in $\phi$, we create one clause vertex $c$.
\item We add one additional super vertex $v$.
\end{itemize}
Next, we describe the edge set of $G$.
\begin{itemize}
\item For each variable $x_i$ in $\phi$ we add the following five edges: 
$x_i x_i^T, x_i x_i^F, x_i^T x_i^F, x_i^T v$ and $ x_i^Fv$.
\item For each pair of variables $x_i,x_j$ in $\phi$ with $i \neq j$ we add the following four edges: 
$x_i^T x_j^T, x_i^Tx_j^F, x_i^Fx_j^T$ and $x_i^Fx_j^F$.
\item For each clause $c$ in $\phi$ we add one edge for each literal. Let $x_i$ appear in $c$. If $x_i$ appears non-negated in $c$ we add edge $c x_i^T$. 
If $x_i$ appears negated in $c$ we add edge $c x_i^F$.
\end{itemize}
This finishes the construction of $G$.
For an illustration see~\cref{fig:NP-example}.

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(x_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=2 \Delta$, $d(c,x_i)=\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a satisfying assignment for $\phi$ as follows. For each variable $x_i$, 
if $\lambda(x_i x_i^T)=\lambda(x_i^T v)$, then we set $x_i$ to \textsc{true}, otherwise we set $x_i$ to \textsc{false}.

@@

Let $\lambda(x_i x_i^T)=t$ and $\lambda(x_i x_i^F)=t'$, for an arbitrary variable $x_i$. 
If both $\lambda(x_i^T v)\neq t+1$ and $\lambda(x_i^F v)\neq t'+1$, then $d(x_i,v)>2$, which is a contradiction. 
Thus, for every variable $x_i$, we have that $\lambda(x_i^T v)= t+1$ or $\lambda(x_i^F v)= t'+1$ (or both). 
In particular, this means that if $\lambda(x_i x_i^F)=\lambda(x_i^F v)$, then we set $x_i$ to \textsc{false}, since in this case $\lambda(x_i x_i^T)\neq\lambda(x_i^T v)$.

@@

Now assume for a contradiction that the described assignment is not satisfying. Then there exists a clause $c$ that is not satisfied. 
Suppose that $x_1, x_2, x_3$ are three variables that appear in $c$.
Recall that we require $d(c,v)=2$ and $d(v,c)=\Delta -1$. 
The fact that $d(c,v)=2$ implies that we must have a temporal path consisting of two edges from $c$ to $v$, 
such that the two edges have consecutive labels. 
By construction of $G$ there are three candidates for such a path, one for each literal of $c$. 
Assume w.l.o.g.\ that $x_1$ appears in $c$ non-negated (the case of a negated appearance of $x_1$ is symmetrical) and that the temporal path realizing $d(c,v)=2$ goes through vertex $x_1^T$. 
Let us denote with $t = \lambda(x_1^T v)$.
It follows that $\lambda(x_1^T c)=\lambda(x_1^T v)-1 = t - 1$.
Furthermore, since $d(c,x_1)=2$ we also have that $\lambda(x_1^T c)=\lambda(x_1 x_1^T)-1$. 
Therefore $\lambda(x_1 x_1^T)=\lambda(x_1^T v) = t$. 
Which implies that $x_1$ is set to \textsc{true}.
%
Let us observe paths from $v$ to $c_1$.
We know that $d(v,c)=\Delta -1$.
The underlying path of the fastest temporal path from $v$ to $c$, that goes through $x_1^T$ is the path $P = (v,x_1^T,c)$.
Since $\lambda(x_1^Tc) > \lambda(x_1^T v)$ we get that the duration of the temporal path $(P,\lambda)$ is equal to 
$d(P,\lambda)= (\Delta + t-1) - t + 1 = \Delta$.
This implies that
the fastest temporal path from $v$ to $c$ is not $(P,\lambda)$ and therefore does not pass through $x_1^T$.
Since there are only two other vertices connected to $c$, 
we have only two other edges incident to $c$, that can be used on a fastest temporal path
$v$ to $c$.
Suppose now w.l.o.g.\ that also $x_2$ appears in $c$ non-negated (the case of a negated appearance of $x_2$ is symmetrical) and that the temporal path realizing $d(v,c)=\Delta-1$ goes through vertex $x_2^T$.
Let us denote with $t' = \lambda(x_2^T v)$.
Since the fastest temporal path from $v$ to $c$ is of the duration $\Delta - 1$,
and the edge $x_2^T c$ is the only edge incident to vertex $c$ and edge $ x_2^T v$,
it follows that $\lambda(x_2^T c) \geq \lambda(x_2^T v) - 2 = t' - 2$.
Since $d(x_2,v) = 2$ it follows that $\lambda(x_2 x_2^T) = \lambda (x_2^T v) - 1 = t' - 1$.
Knowing this and the fact that $d(x_2,c)=2$, we get that $\lambda(x_2^T c)$ must be equal to $t'-2$.
Therefore the fastest temporal path from $v$ to $c$ passes through edges $ x_2^Tv$ and $x_2^Tc$.
In the above we have also determined that $\lambda(x_2x_2^T) \neq \lambda ( x_2^Tv)$,
which implies that $x_2$ is set to \textsc{false}.
But now we have that $x_1,x_2$ both appear in $c$ non-negated, where one of them is \textsc{true}, while the other is \textsc{false},
which implies that the clause $c$ is satisfied, a contradiction.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{true}, we set $\lambda(x_i^F v)=3$.
    \item If variable $x_i$ is set to \textsc{false}, we set $\lambda(x_i^T v)=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all duratios are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

    If $x_i$ is set to \textsc{true}, then there is a temporal path from $x_i$ to $v$ via 
    $x_i^F$ of duration $2$, since
    $\lambda(x_i x_i^F)=2$ and $\lambda(x_i^F v)=3$.
    For a temporal path from $v$ to $x_i$ we observe the following.
    The only possible labels to leave the vertex $v$ are $2$ and $3$, which take us from $v$ to $x_j^T$ or $x_j^F$ of some variable $x_j$.
    The only two edges incident to $x_i$ have labels $2$, therefore the fastest path from $v$ to $x_i$
    cannot finish before the time $\Delta + 2$.
    The fastest way to leave $v$ and enter to $x_i$ would then be to leave $v$ at edge $x_i^Fv$ with label $3$,
    and continue to $x_i$ at time $\Delta + 2$,
    which gives us the desired duration $\Delta$.

@@

    Suppose $x_i,x_j,x_k$ appear in $c$.
    Since we have a satisfying assignment at least one of the literals in $c$ is set to \textsc{true} and at least one to \textsc{false}. 
    Suppose $x_i$ is the variable of the literal that is \textsc{true} in $c$,
    and $x_j$ is the variable of the literal that is \textsc{false} in $c$.
    Let $x_i$ appear non-negated in $c$ and is therefore set to \textsc{true} (the case when $x_i$ appears negated in $c$ and is set to \textsc{false} is symmetric).
    Then there is a temporal path from $c$ to $v$ through $x_i^T$ such that $\lambda(x_i^T c)=1$ and $\lambda(x_i^T v)=2$. 
    Let $x_j$ appear non-negated in $c$ and is therefore set to \textsc{false} 
    (the case when $x_j$ appears negated in $c$ and is set to \textsc{true} is symmetric).
    Then there is a temporal path from $v$ to $c$ through $x_j^T$ such that $\lambda(x_j^T v)=3$ and $\lambda(x_j^T c)=1$,
    which results in a temporal path from $v$ to $c$ of duration $\Delta -1$.

@@

    There is a temporal path from $c$ to $x_i$ via $x_i^T$ and also a temporal path from $c$ to $x_i^F$ via $x_i^T$ such that $\lambda(x_i^T c)=1$ and $\lambda(x_i x_i^T)=\lambda(x_i^T x_i^F)=2$,
    which proves the first equality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^T$ and
    second, from $x_i^F$ to $c$ through $x_i^T$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(x_i x_i^T) = \lambda (x_i^T x_i^F)=2$ and 
    finish on the edge with label $1$, as $\lambda(x_i^T c) = 1$.

@@

    There is a temporal path from $c$ to $x$ via $x^F$ and also a temporal path from $c$ to $x^T$ via $x^F$ such that $\lambda(c x^F)=1$ and $\lambda(x x^F)=\lambda(x^T x^F)=2$,
    which proves the first inequality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^F$ and
    second, from $x_i^T$ to $c$ through $x_i^F$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(x_i x_i^F) = \lambda (x_i^T x_i^F)=2$ and 
    finish on the edge with label $1$, as $\lambda(x_i^F c) = 1$.
    Which proves the second equality.

@@

    Let $x_j$ be a variable that appears non-negated in $c$ (the case where $x_j$ appears negated is symmetric). 
    Then there is a temporal path 
    from $c$ to $x_i^T$ via $x_j^T$ and also a temporal path 
    from $c$ to $x_i^F$ via $x_j^T$ such that $\lambda(x_j^T c)=1$ and $\lambda(x_j^T x_i^T)=\lambda(x_j^T x_i^F)=2$,
    which proves the first equality.
    Using the same temporal path in the opposite direction,
    \ie first the edge $x_j^T c$ and then one of the edges $x_j^T x_i^F$ or $x_j^T x_i^T$ at times $2$ and $\Delta + 1$, respectively,
    yields the second equality.
    For a temporal path from $c$ to $x_i$ we traverse the following three edges 
    $x_j^T c, x_j^T x_i^F$ and $x_i^Fx_i$,
    with labels $1, 2, 2$ respectively (\ie the path traverses them at time $1,2$ and $\Delta + 2$, respectively), which proves the third equality.
    Now for the case of a temporal path from $x_i$ to $c$,
    we use the same three edges, but in the opposite direction,
    namely $x_i^Fx_i, x_j^T x_i^F$ and $x_j^T c$,
    again at times $2, \Delta + 2, 2\Delta + 1$, respectively,
    which proves the last equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $x_k^T c_i, x_\ell ^T c_j, x_k^T x_\ell ^T$, with labels $1,2,1$ (at times $1, 2, \delta + 1$), respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, therefore $\Delta\ge 3$.
\end{proof}

@@

In the case when we want to guess the fastest path from $u$, that (somehow) passes through the vertex $x$, and goes directly to $v$ via an edge, we write it as a fastest path of form $u \leadsto x \rightarrow v$.
Similarly it holds if we want to include more vertices that have to be visited by the path. If there is an edge (\ie a unique fastest path) among two vertices, we denote it by $\rightarrow$,
if the fastest path among two vertices is not uniquely determined, we denote it by $\leadsto$.

@@

    The fastest temporal paths among all pairs of vertices of $U$.

>>>>

\begin{abstract}
In this paper we initiate the study of the \emph{temporal graph realization} problem with respect to the fastest path durations between its vertices, 
while we focus on periodic temporal graphs. 
Given an $n \times n$ matrix $D$ and a $\Delta \in \mathbb{N}$, the goal is to construct a $\Delta$-periodic temporal graph with $n$ vertices 
such that the duration of a \emph{fastest path} from $v_i$ to $v_j$ is equal to $D_{i,j}$, or to decide that such a temporal graph does not exist. 
The variations of the problem on static graphs has been well studied and understood since the 1960's (e.g.\ [Erd\H{o}s and Gallai, 1960], [Hakimi and Yau, 1965]), and this area of research remains active until nowadays.

@@

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity between all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all these cases, the duration of the temporal paths has not been considered.

@@

We remark that a fastest path between two vertices in a temporal graph can be computed in polynomial time~\cite{xuan_computing_2003,Wu2016Efficient}.
%Wu et~al.~\cite{Wu2016Efficient} provide an algorithm that computes the duration of the fastest temporal path from a source vertex $v$, in a temporal graph $(G,\lambda)$, to all other vertices,
%that runs in polynomial-time.
Hence, given a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, we can compute in polynomial-time  the matrix $D$
which consists of durations of fastest temporal paths between all pairs of vertices in $(G,\lambda,\Delta)$.

@@

 We start by describing the vertex set of the underlying graph $G$ of $D$.
\begin{itemize}
\item For each variable $x_i$ in $\phi$, we create three variable vertices $x_i, x_i^T, x_i^F$.
\item For each clause $c$ in $\phi$, we create one clause vertex $c$.
\item We add one additional super vertex $v$.
\end{itemize}
Next, we describe the edge set of $G$.
\begin{itemize}
\item For each variable $x_i$ in $\phi$ we add the following five edges: 
$\{x_i, x_i^T\}$, $\{x_i, x_i^F\}$, $\{x_i^T, x_i^F\}$, $\{x_i^T, v\}$, and $\{x_i^F,v\}$.
\item For each pair of variables $x_i,x_j$ in $\phi$ with $i \neq j$ we add the following four edges: 
$\{x_i^T, x_j^T\}$, $\{x_i^T,x_j^F\}$, $\{x_i^F,x_j^T\}$, and $\{x_i^F,x_j^F\}$.
\item For each clause $c$ in $\phi$ we add one edge for each literal. Let $x_i$ appear in $c$. If $x_i$ appears non-negated in $c$ we add edge $\{c, x_i^T\}$. 
If $x_i$ appears negated in $c$ we add edge $\{c, x_i^F\}$.
\end{itemize}
This finishes the construction of $G$.
For an illustration see~\cref{fig:NP-example}.

%
\begin{figure}[t]
	\noindent
	\makebox[\textwidth]{
		\centering
	\includegraphics{fig-NP-reduction}
}
\caption{Illustration of the temporal graph $(G,\lambda)$ from the NP-hardness reduction, 
	where the NAE 3-SAT formula $\phi$ is of the form $\phi = \text{NAE}(x_1, \overline{x}_2, x_3) \wedge \text{NAE}(x_1, x_2, x_4)$.
To improve the readability, we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neq j$) with gray dotted lines.
Presented is the labeling of $G$ corresponding to the assignment $x_1=x_2=\textsc{true}$ and $x_3,x_4=\textsc{false}$,
where all unlabeled edges get the label $2$.
}\label{fig:NP-example}
\end{figure}
%
We set $\Delta$ to some constant larger than two, that is, $\Delta\ge 3$. Next, we specify the durations in the matrix $D$ between all vertex pairs.
For the sake of simplicity we write $D_{u,v}$ as $d(u,v)$,
where $u,v$ are two vertices of $G$. 
We start by setting the value of $d(u,v) = 1$ where $u$ and $v$ are two adjacent vertices in $G$.
\begin{itemize}
    %\item For each variable $x$ in $\phi$ we specify the following distances between the non-adjacent variable vertices:
    %$d(x_1,x_2)=2$. 
    %\item For pair of variable $x,y$ in $\phi$ with $x\neq y$ we specify the following 
    %$d(x_1,y^T)=d(x_1,y^F)=3$, $d(x_2,y^T)=d(x_2,y^F)=2$.
    \item For each variable $x_i$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(x_i,v)=2$ and $d(v,x_i)= \Delta $. %, $d(x_2,v)=3$.
    \item For each clause $c$ in $\phi$ and the super vertex $v$
    we specify the following durations:
    $d(c,v)=2$ and $d(v,c)= \Delta - 1$.
    \item Let $x_i$ be a variable that appears in clause $c$, then  we specify the following durations:
    $d(c,x_i)=2$ and $d(x_i,c)=\Delta$.
    If $x_i$ appears non-negated in $c$ we specify the following durations:
    $d(c,x_i^F)=2$ and $d(x_i^F,c)=\Delta$.
    If $x_i$ appears negated in $c$ we specify the following duratios:
    $d(c,x_i^T)=2$ and $d(x_i^T,c)=\Delta$.
    \item Let $x_i$ be a variable that does \emph{not} appear in clause $c$, then we specify the following duratios:
    $d(x_i,c)=2 \Delta$, $d(c,x_i)=\Delta + 2$
    and
    $d(c,x_i^T)=d(c,x_i^F)=2$, $d(x_i^T,c)=d(x_i^F,c)=\Delta$. 
    \item For each pair of variables $x_i \neq x_j$ in $\phi$ we specify the following duratios:
    $d(x_i,x_j)=2\Delta +1$ and
    $d(x_i,x_j^T)=d(x_i,x_j^F)=\Delta + 1$.
    \item For each pair of clauses $c_i \neq c_j$ in $\phi$ we specify the following duratios:
    $d(c_i,c_j)= \Delta + 1$.
\end{itemize}
This finishes the construction of the instance $(D,\Delta)$ of \deltaExact\, which can clearly be done in polynomial time. In the remainder we show that $(D,\Delta)$ is a \textsc{Yes}-instance of \deltaExact\ if and only if NAE 3-SAT formula $\phi$ is satisfiable.

@@

We construct a satisfying assignment for $\phi$ as follows. For each variable $x_i$, 
if $\lambda(\{x_i, x_i^T\})=\lambda(\{x_i^T, v\})$, then we set $x_i$ to \textsc{true}, otherwise we set $x_i$ to \textsc{false}.

@@

Let $\lambda(\{x_i, x_i^T\})=t$ and $\lambda(\{x_i, x_i^F\})=t'$, for an arbitrary variable $x_i$. 
If both $\lambda(\{x_i^T, v\})\neq t+1$ and $\lambda(\{x_i^F, v\})\neq t'+1$, then $d(x_i,v)>2$, which is a contradiction. 
Thus, for every variable $x_i$, we have that $\lambda(\{x_i^T, v\})= t+1$ or $\lambda(\{x_i^F, v\})= t'+1$ (or both). 
In particular, this means that if $\lambda(\{x_i, x_i^F\})=\lambda(\{x_i^F, v\})$, then we set $x_i$ to \textsc{false}, since in this case $\lambda(\{x_i, x_i^T\})\neq\lambda(\{x_i^T, v\})$.

@@

Now assume for a contradiction that the described assignment is not satisfying. Then there exists a clause $c$ that is not satisfied. 
Suppose that $x_1, x_2, x_3$ are three variables that appear in $c$.
Recall that we require $d(c,v)=2$ and $d(v,c)=\Delta -1$. 
The fact that $d(c,v)=2$ implies that we must have a temporal path consisting of two edges from $c$ to $v$, 
such that the two edges have consecutive labels. 
By construction of $G$ there are three candidates for such a path, one for each literal of $c$. 
Assume w.l.o.g.\ that $x_1$ appears in $c$ non-negated (the case of a negated appearance of $x_1$ is symmetrical) and that the temporal path realizing $d(c,v)=2$ goes through vertex $x_1^T$. 
Let us denote with $t = \lambda(\{x_1^T, v\})$.
It follows that $\lambda(\{x_1^T, c\})=\lambda(\{x_1^T, v\})-1 = t - 1$.
Furthermore, since $d(c,x_1)=2$ we also have that $\lambda(\{x_1^T, c\})=\lambda(\{x_1, x_1^T\})-1$. 
Therefore $\lambda(\{x_1, x_1^T\})=\lambda(\{x_1^T, v\}) = t$. 
Which implies that $x_1$ is set to \textsc{true}.
%
Let us observe paths from $v$ to $c$.
We know that $d(v,c)=\Delta -1$.
The underlying path of the fastest temporal path from $v$ to $c$, that goes through $x_1^T$ is the path $P = (v,x_1^T,c)$.
Since $\lambda(\{x_1^T,c\}) > \lambda(\{x_1^T, v\})$ we get that the duration of the temporal path $(P,\lambda)$ is equal to 
$d(P,\lambda)= (\Delta + t-1) - t + 1 = \Delta$.
This implies that
the fastest temporal path from $v$ to $c$ is not $(P,\lambda)$ and therefore does not pass through $x_1^T$.
Since there are only two other vertices connected to $c$, 
we have only two other edges incident to $c$, that can be used on a fastest temporal path
$v$ to $c$.
Suppose now w.l.o.g.\ that also $x_2$ appears in $c$ non-negated (the case of a negated appearance of $x_2$ is symmetrical) and that the temporal path realizing $d(v,c)=\Delta-1$ goes through vertex $x_2^T$.
Let us denote with $t' = \lambda(\{x_2^T, v\})$.
Since the fastest temporal path from $v$ to $c$ is of the duration $\Delta - 1$,
and the edge $x_2^T c$ is the only edge incident to vertex $c$ and edge $\{x_2^T, v\}$,
it follows that $\lambda(\{x_2^T, c\}) \geq \lambda(\{x_2^T, v\}) - 2 = t' - 2$.
Since $d(x_2,v) = 2$ it follows that $\lambda(\{x_2, x_2^T\}) = \lambda (\{x_2^T, v\}) - 1 = t' - 1$.
Knowing this and the fact that $d(x_2,c)=2$, we get that $\lambda(\{x_2^T, c\})$ must be equal to $t'-2$.
Therefore the fastest temporal path from $v$ to $c$ passes through edges $\{x_2^T,v\}$ and $\{x_2^T,c\}$.
In the above we have also determined that $\lambda(\{x_2,x_2^T\}) \neq \lambda ( \{x_2^T,v\})$,
which implies that $x_2$ is set to \textsc{false}.
But now we have that $x_1,x_2$ both appear in $c$ non-negated, where one of them is \textsc{true}, while the other is \textsc{false},
which implies that the clause $c$ is satisfied, a contradiction.

@@

We construct a labeling $\lambda$ as follows.
\begin{itemize}
    \item All edges incident with a clause vertex $c$ obtain label one.
    %\item For each variable $x$, we set $\lambda(\{x^T,x^F\})=1$.
    \item If variable $x_i$ is set to \textsc{true}, we set $\lambda(\{x_i^F, v\})=3$.
    \item If variable $x_i$ is set to \textsc{false}, we set $\lambda(\{x_i^T, v\})=3$.
    \item We set the labels of all other edges to two.
\end{itemize}
For an example of the constructed temporal graph see~\cref{fig:NP-example}.
We now verify that all duratios are realized.
\begin{itemize}
    \item For each variable $x_i$ in $\phi$ we have to check that $d(x_i,v)=2$
    and $d(v,x_i)=\Delta$.

    If $x_i$ is set to \textsc{true}, then there is a temporal path from $x_i$ to $v$ via 
    $x_i^F$ of duration $2$, since
    $\lambda(\{x_i, x_i^F\})=2$ and $\lambda(\{x_i^F, v\})=3$.
    For a temporal path from $v$ to $x_i$ we observe the following.
    The only possible labels to leave the vertex $v$ are $2$ and $3$, which take us from $v$ to $x_j^T$ or $x_j^F$ of some variable $x_j$.
    The only two edges incident to $x_i$ have labels $2$, therefore the fastest path from $v$ to $x_i$
    cannot finish before the time $\Delta + 2$.
    The fastest way to leave $v$ and enter to $x_i$ would then be to leave $v$ at edge $\{x_i^F,v\}$ with label $3$,
    and continue to $x_i$ at time $\Delta + 2$,
    which gives us the desired duration $\Delta$.

@@

    Suppose $x_i,x_j,x_k$ appear in $c$.
    Since we have a satisfying assignment at least one of the literals in $c$ is set to \textsc{true} and at least one to \textsc{false}. 
    Suppose $x_i$ is the variable of the literal that is \textsc{true} in $c$,
    and $x_j$ is the variable of the literal that is \textsc{false} in $c$.
    Let $x_i$ appear non-negated in $c$ and is therefore set to \textsc{true} (the case when $x_i$ appears negated in $c$ and is set to \textsc{false} is symmetric).
    Then there is a temporal path from $c$ to $v$ through $x_i^T$ such that $\lambda(\{x_i^T, c\})=1$ and $\lambda(\{x_i^T, v\})=2$. 
    Let $x_j$ appear non-negated in $c$ and is therefore set to \textsc{false} 
    (the case when $x_j$ appears negated in $c$ and is set to \textsc{true} is symmetric).
    Then there is a temporal path from $v$ to $c$ through $x_j^T$ such that $\lambda(\{x_j^T, v\})=3$ and $\lambda(\{x_j^T, c\})=1$,
    which results in a temporal path from $v$ to $c$ of duration $\Delta -1$.

@@

    There is a temporal path from $c$ to $x_i$ via $x_i^T$ and also a temporal path from $c$ to $x_i^F$ via $x_i^T$ such that $\lambda(\{x_i^T, c\})=1$ and $\lambda(\{x_i, x_i^T\})=\lambda(\{x_i^T, x_i^F\})=2$,
    which proves the first equality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^T$ and
    second, from $x_i^F$ to $c$ through $x_i^T$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(\{x_i, x_i^T\}) = \lambda (\{x_i^T, x_i^F\})=2$ and 
    finish on the edge with label $1$, as $\lambda(\{x_i^T, c\}) = 1$.

@@

    There is a temporal path from $c$ to $x$ via $x^F$ and also a temporal path from $c$ to $x^T$ via $x^F$ such that $\lambda(\{c, x^F\})=1$ and $\lambda(\{x, x^F\})=\lambda(\{x^T, x^F\})=2$,
    which proves the first inequality.
    There are also the following two temporal paths,
    first, from $x_i$ to $c$ through $x_i^F$ and
    second, from $x_i^T$ to $c$ through $x_i^F$.
    Both of the temporal paths start on the edge with label $2$, as 
    $\lambda(\{x_i, x_i^F\}) = \lambda (\{x_i^T, x_i^F\})=2$ and 
    finish on the edge with label $1$, as $\lambda(\{x_i^F, c\}) = 1$.
    Which proves the second equality.

@@

    Let $x_j$ be a variable that appears non-negated in $c$ (the case where $x_j$ appears negated is symmetric). 
    Then there is a temporal path 
    from $c$ to $x_i^T$ via $x_j^T$ and also a temporal path 
    from $c$ to $x_i^F$ via $x_j^T$ such that $\lambda(\{x_j^T, c\})=1$ and $\lambda(\{x_j^T, x_i^T\})=\lambda(\{x_j^T, x_i^F\})=2$,
    which proves the first equality.
    Using the same temporal path in the opposite direction,
    \ie first the edge $x_j^T c$ and then one of the edges $\{x_j^T, x_i^F\}$ or $\{x_j^T, x_i^T\}$ at times $2$ and $\Delta + 1$, respectively,
    yields the second equality.
    For a temporal path from $c$ to $x_i$ we traverse the following three edges 
    $\{x_j^T, c\}$, $\{x_j^T, x_i^F\}$, and $\{x_i^F,x_i\}$,
    with labels $1$, $2$, and $2$ respectively (\ie the path traverses them at time $1,2$ and $\Delta + 2$, respectively), which proves the third equality.
    Now for the case of a temporal path from $x_i$ to $c$,
    we use the same three edges, but in the opposite direction,
    namely $\{x_i^F,x_i\}$, $\{x_j^T, x_i^F\}$, and $\{x_j^T, c\}$,
    again at times $2$, $\Delta + 2$, and $2\Delta + 1$, respectively,
    which proves the last equality.
    Note that all of the above temporal paths are also the shortest possible, 
    and since the labels of first and last edges (of these paths) are unique,
    it follows that we cannot find faster temporal paths.

@@

    Let $x_k$ be a variable that appears non-negated in $c_i$
    and $x_\ell$ the variable that appears non-negated in $c_j$
    (all other cases are symmetric).
    There is a path of length three from $c_i$ to $c_j$
    that passes first through vertex $x_k^T$ and then through vertex $x_\ell^T$.
    Therefore the temporal path from $c_i$ to $c_j$
    uses the edges $\{x_k^T, c_i\}$, $\{x_\ell^T, c_j\}$, and $\{x_k^T, x_\ell ^T\}$, with labels $1$,~$2$, and $1$ (at times $1$,~$2$, and $\delta + 1$), respectively,
    which proves the desired equality.
    Note also that this is the shortest path between $c_i$ and $c_j$, and that the first and the last edge must have the label $1$,
    therefore it follows that this is the fastest temporal path.
\end{itemize}
Lastly, observe that the above constructed labeling $\lambda$ uses values $\{1,2,3\}\subseteq[\Delta]$, therefore $\Delta\ge 3$.
\end{proof}

@@

In the case when we want to guess the fastest path from $u$, that (somehow) passes through the vertex $x$, and goes directly to $v$ via an edge, we write it as a fastest path of the form $u \leadsto x \rightarrow v$.
Similarly it holds if we want to include more vertices that have to be visited by the path. If there is an edge (\ie a unique fastest path) between two vertices, we denote it by $\rightarrow$,
if the fastest path between two vertices is not uniquely determined, we denote it by $\leadsto$.

@@

    The fastest temporal paths between all pairs of vertices of $U$.
